---
title: "GSS challange Less Data"
author: "Edward Gomez"
date: "October 2025"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

White people approve of the capital punishment at a ratio of about 2:1. Black people have about an even amount of people who oppose or approve of capital punishment. The other group has a higher proportion of people who approve of capital punishment when compared to black people, but lower than white people. 

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table2 <- xtabs(~cappun + race, data=mydata)
rowPerc(table2)
colPerc(table2)
```

A majority of white people, 67.75%, favor capital punishment, while 32.25% of them oppose it. A minority of black people, 47.24%, favor capital punishment, while 57.26% of them oppose it. A majority of other people, 61.25%, favor capital punishment, while 38.75% of them oppose it.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
fisherXY(mydata$cappun,mydata$race,printAll = TRUE)
```

The p-value found between the "white" and "black" groups was 0.0000741 which shows they are statistically different. The p-value found between the "white" and "other" groups was 0.2565 which shows they are not statistically different.The p-value found between the "other" and "black" groups was 0.06299 which shows they are not statistically different.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In the beginning I decided this was a categorical~categorical problem so I used the corresponding file to help me solve the problem. Graphing the data I found that white people had a greater proportion of people who approved of capital punishment than the "black" and "other" groups did. When I put the data in a table I found that a majority of white people, 67.75%, favor capital punishment, while 32.25% of them oppose it, a minority of black people, 47.24%, favor capital punishment, while 57.26% of them oppose it, and a majority of other people, 61.25%, favor capital punishment, while 38.75% of them oppose it. When I ran a fisher XY test, I found that there was only a statistical difference between the "black" and "white" groups. This is in stark contrast to my original findings that showed there was a statistical difference between all the groups. This difference is due to the fact my original findings were made using more data. 
